---
title: "Sentiment Analysis using a Word Cloud"
author: "Riki Saito"
date: "August 7, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
pacman::p_load(dplyr, stringr, rvest, wordcloud, tm, colorspace, slam)




#Remove all white space
trim <- function (x) gsub("^\\s+|\\s+$", "", x)

#Clean strings (lower case, remove punctation, remove all white space)
str_clean <- function(strings) {
  require(dplyr)
  strings %>%
    tolower() %>%
    removePunctuation() %>%
    stripWhitespace() %>%
    trim()  
}



#wrapper function for sentiment score, from https://www.r-bloggers.com/sentiment-analysis-on-donald-trump-using-r-and-tableau/
sentiment_scores = function(tweets, positive_words, negative_words){
  scores = sapply(tweets,
                 function(tweet, positive_words, negative_words){
                   tweet = gsub("[[:punct:]]", "", tweet)    # remove punctuation
                   tweet = gsub("[[:cntrl:]]", "", tweet)   # remove control characters
                   #tweet = gsub('\\d+', '', tweet)          # remove digits
                   
                   # Let's have error handling function when trying tolower
                   tryTolower = function(x){
                     # create missing value
                     y = NA
                     # tryCatch error
                     try_error = tryCatch(tolower(x), error=function(e) e)
                     # if not an error
                     if (!inherits(try_error, "error"))
                       y = tolower(x)
                     # result
                     return(y)
                   }
                   # use tryTolower with sapply
                   tweet = sapply(tweet, tryTolower)
                   # split sentence into words with str_split function from stringr package
                   word_list = str_split(tweet, "\\s+")
                   words = unlist(word_list)
                   # compare words to the dictionaries of positive & negative terms
                   positive.matches = match(words, positives)
                   negative.matches = match(words, negatives)
                   # get the position of the matched term or NA
                   # we just want a TRUE/FALSE
                   positive_matches = !is.na(positive.matches)
                   negative_matches = !is.na(negative.matches)
                   # final score
                   score = sum(positive_matches) - sum(negative_matches)
                   return(score)
                 }, positive_matches, negative_matches)
  return(scores)
}

```


<!-- #load source functions -->
<!-- source("https://raw.githubusercontent.com/rjsaito/Just-R-Things/master/Natural%20Language%20Processing/str_cleaner.R") -->
<!-- source("https://raw.githubusercontent.com/rjsaito/Just-R-Things/master/Natural%20Language%20Processing/Sentiment%20Analysis/sentiment_scores.R") -->

#load reviews
reviews <- read.csv("https://raw.githubusercontent.com/rjsaito/Just-R-Things/master/Natural%20Language%20Processing/Sentiment%20Analysis/reviews_ueboom.csv", stringsAsFactors = F)


#load lexicon
positives <- readLines("https://raw.githubusercontent.com/rjsaito/Just-R-Things/master/Natural%20Language%20Processing/lexicon/positive-words.txt")
negatives <- readLines("https://raw.githubusercontent.com/rjsaito/Just-R-Things/master/Natural%20Language%20Processing/lexicon/negative-words.txt")


reviews_clean <- reviews$comment %>%
  str_clean() %>%
  VectorSource() %>%
  Corpus() %>%
  tm_map(PlainTextDocument) %>%
  tm_map(removeWords, stopwords('english')) 

#sentiment scores
review_score <- sentiment_scores(reviews_clean, positives, negatives)
names(review_score) = 1:length(review_score)

reviews_tm <- TermDocumentMatrix(reviews_clean)
word_score <- with(reviews_tm, data.frame(doc = j, word_ind = i)) %>%
  merge(data.frame(review_score), by.x = "doc", by.y = "row.names") %>%
  group_by(word_ind) %>%
  summarise(avg_sc = mean(review_score)) %>%
  merge(data.frame(word = reviews_tm$dimnames$Terms, stringsAsFactors = F), by.x = "word_ind", by.y = "row.names")

word_color <- word_score$avg_sc %>%
  rank() %>%
  round() %>%
  diverge_hcl(diff(range(.)), h = c(0, 120))[.]


freq <- row_sums(reviews_tm)
words <- names(freq)

wordcloud(words, freq, colors = word_color, max.words = 100)  
```
